% ============================================================
% chapter_06_conclusion.tex - Chapter 6, Conclusion
%
% Purpose: summarises the empirical findings, the methodological
%          finding (full reproducibility on free cloud infrastructure),
%          and the recommendations for the wider research community.
% Sections: 6.1 Summary of Findings, 6.2 Implications,
%           6.3 Recommendations.
% Included by main_thesis.tex.
% License: CC BY 4.0 (see ../LICENSE).
% ============================================================
%
% Source material:
%   - paper/PPS1691_..._compact_20260526.tex  (conclusion)
%   - paper/thesis_outline.md                  (§6.1-§6.3 mapping)
%
% Conventions: no em-dashes; no AI-language; no overclaim;
%              dual-contribution framing (scientific + reproducibility).
% ============================================================

\chapter{Conclusion}\label{chap:conclusion}

This thesis asked how far a marginal-correction bias correction
framework can take a daily satellite precipitation field, and it
answered the question in two registers: what the framework
delivers as a scientific result, and what it delivers as a
reproducible methodological artefact. This chapter summarises
the findings, states their implications, and sets out
recommendations for users, researchers, and the institutions that
maintain the underlying data.


% ============================================================
\section{Summary of Findings}\label{sec:conclusion-summary}
% ============================================================

The scientific finding is a stage-resolved account of what the
LSEQM+DL (Linear Scaling, Empirical Quantile Mapping, Generalized
Pareto Distribution tail correction, and Deep Learning refinement
through a Convolutional Neural Network) pipeline can and cannot
do. Applied to the Integrated Multi-satellitE Retrievals for
Global Precipitation Measurement (GPM) Late Run (IMERG-L) over
Indonesia for 2001 to 2025 and validated against 171 independent
Meteorological, Climatological, and Geophysical Agency (BMKG)
stations, the pipeline moves three of the four verification
pillars cleanly toward the gauge target. The value-adjustment
pillar reaches a relative bias near zero and a standard deviation
ratio of unity; the distribution-alignment pillar reaches a
Kolmogorov-Smirnov $p$-value at which the corrected and gauge
distributions are no longer statistically distinguishable; and
the extreme-preservation pillar reaches $Q_{99}$ within one
percent of the gauge target. The fourth pillar, event detection,
shows a designed trade-off: the aggregate probability of
detection falls as the dry-day-matching mechanism suppresses
spurious light rain, but the threshold-stratified view shows the
LSEQM+DL pipeline overtaking Linear Scaling above the
$20$~mm/day threshold that drives flood and drought decisions.
Each stage moves the metrics matched to its design dimension and
leaves the others alone.

The result that gives the thesis its title is the one metric that
does not move. The Pearson correlation between paired daily
satellite and gauge values stays near $0.34$ across all three
corrected stages, a shift smaller than the single-pixel sampling
uncertainty, and the same flat pattern holds station-by-station
in the Taylor analysis and month-by-month across the seasonal
cycle. This is not a failure of the implementation. It is the
predicted behaviour of any marginal correction family: rank
preservation and variance inflation without time-dependent skill
together bound the attainable correlation at the day-by-day
pairing quality of the raw satellite retrieval, which lies
upstream of every step in the pipeline. The corrected value of
$0.348$ sits inside the empirical band of $0.3$ to $0.5$ reported
for daily IMERG over tropical regions. The answer to the title
question is therefore precise: a marginal-correction framework
can take a daily field all the way to the gauge marginal
distribution at every percentile, and no distance at all along
the axis of day-by-day timing.

The methodological finding is that this entire account is
independently reproducible. The pipeline is released as a
configuration-driven set of source modules and numbered
notebooks that run unchanged in a local environment or on a
free-tier cloud instance. The end-to-end sequence was re-executed
on a free Google Colab instance for the Bali subdomain across all
36 dekadal windows in 72.1 minutes on a standard Central
Processing Unit (CPU) runtime,
which establishes that an independent reader can reconstruct the
results for a self-contained region without institutional compute
access; the full Indonesian domain scales up with pixel count and
is correspondingly longer. The two findings are of equal
weight: the scientific result is credible because it is
reproducible, and the reproducible artefact is worth releasing
because it carries a result worth re-verifying.


% ============================================================
\section{Implications}\label{sec:conclusion-implications}
% ============================================================

For operational use, the central implication is that the
corrected product should be matched to the metrics that moved.
The corrected daily IMERG-L is suitable for the broad class of
applications that depend on distributional properties:
dekadal-to-monthly water budgets, drought indicators such as the
Standardized Precipitation Index and its evapotranspiration
variant, extreme-event frequency analysis, threshold-exceedance
climatology, and hydrological model calibration against multi-year
statistics. It is not suitable, on its own, for applications that
depend on the precise calendar day of an event: real-time
flood-event nowcasting, single-event forecasting, and day-1
hydrological initialisation inherit the timing ceiling of the raw
retrieval and must combine the corrected product with sub-daily
disaggregation, move to ensemble methods, or accept the timing
limitation. Correcting at the daily scale remains necessary so
that the dekadal, monthly, and seasonal aggregates derived from
the daily field are bias-free.

For the open-science contribution, the implication is that the
verification posture argued for in this thesis can be applied by
others. The four-pillar framework with Pearson correlation
reported as a separate temporal-skill diagnostic, rather than
folded into a single skill score, lets a structural limit and a
methodological improvement appear in the same report. An
independent group can apply the same verification scaffolding to
a different product, a different reference, or a different region,
and can reproduce or challenge the present result on its own
terms.


% ============================================================
\section{Recommendations}\label{sec:conclusion-recommendations}
% ============================================================

\medskip\noindent\textbf{For users of corrected satellite
precipitation.}
Read the verification report before choosing an application.
A product that matches the gauge distribution is the right input
for distributional questions and the wrong input for
day-specific timing questions, and the two cases cannot be told
apart from an aggregate skill score alone. Where the application
is timing-sensitive and operates over the Maritime Continent, the
UTC-aligned daily window is a known contributor to the timing
ceiling, and a locally-aligned re-aggregation should be
considered before the product is rejected as unskilful.

\medskip\noindent\textbf{For researchers extending the method.}
The temporal-skill ceiling is set by the raw retrieval, not by
the post-processing, so raising it requires methods outside the
marginal correction family: multivariate quantile mapping,
sub-daily disaggregation against the native half-hourly product,
stochastic disaggregation with ensemble verification, or
local-day re-aggregation with a geostationary diurnal-cycle
reference. The three sensitivity parameters explored one-at-a-time in
Section~\ref{sec:results-sensitivity} but not jointly optimised
(the blending weight, the GPD threshold, and the density
saturation count) are exposed in the configuration and are the
natural target of a full Cartesian grid search with a strict
temporal hold-out. A transfer experiment to a second tropical
region would test the portability that the framework's design
allows but does not by itself establish.

\medskip\noindent\textbf{For BMKG, CPC, and Institut Pertanian Bogor (IPB).}
The reproducibility contribution depends on continued access to
the underlying archives and on the maintenance of the released
code against dependency drift. A containerised distribution and a
versioned data snapshot would make the reproducibility claim
durable beyond the dependency state recorded at release. More
broadly, the gap this thesis addresses, namely the rarity of
released code in the bias correction literature, is one that
institutional data-sharing and reproducibility standards are well
placed to close, and the framework released here is offered as a
concrete example of what such standards can produce.
